---
title: "Introduction to `aopdata`"
author: "Marcus Saraiva"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
abstract: "This vignette introduces the `aopdata` package, lists the cities for which data are available for download, and presents the data dictionary."
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{Introduction to `aopdata`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


**`aopdata`** is an R package that makes it easy to download data from the 
[Access to Opportunities Project (AOP)](https://www.ipea.gov.br/acessooportunidades/en/). 
AOP is a research initiative led by the Institute for Applied Economic Research (Ipea) 
that aims to study transport accessibility and inequalities in access to opportunities 
in Brazilian cities. You can access more information about the project on the [AOP website](https://www.ipea.gov.br/acessooportunidades/en/).


## Installation

You can install `aopdata` from CRAN, or the development version from GitHub.

```{r, eval = FALSE}
# CRAN
install.packages('aopdata')

# github
devtools::install_github("ipeaGIT/aopdata", subdir = "r-package")
```

## Usage

After installation, you can easily download accessibility estimates, as well as 
population and land use data, from 20 major Brazilian cities. Those cities are:

| **Abbreviation** | **City Name**       |
|--------------|-----------------|
| bel          | Belem           |
| bho          | Belo Horizonte  |
| bsb          | Brasilia        |
| cam          | Campinas        |
| cgr          | Campo Grande    |
| cur          | Curitiba        |
| duq          | Duque de Caxias |
| for          | Fortaleza       |
| goi          | Goiania         |
| gua          | Guarulhos       |
| mac          | Maceio          |
| man          | Manaus          |
| nat          | Natal           |
| poa          | Porto Alegre    |
| rec          | Recife          |
| rio          | Rio de Janeiro  |
| sal          | Salvador        |
| sgo          | Sao Gonçalo     |
| slz          | Sao Luis        |
| spo          | Sao Paulo       |

**Read *accessibility* estimates**

To download accessibility estimates, you can use the `read_access()` function. The
example below downloads data on accessibility by walking for the city of Curitiba, 
in Southern Brazil. The function will download data for the year 2019, which is the 
one currently available.

`read_access()` can also download the spatial geometry of each city by setting
`geometry = TRUE`, which can be used to map accessibility levels across the
city. Refer to the [mapping population and land use](landuse_maps.html) and
[mapping urban accessibility](access_maps.html) vignettes for examples of how
to use the spatial geometry information. For now, we'll set `geometry = FALSE` to 
download a dataframe with no spatial information. 

```{r include=FALSE}
# load the aopdata package
library(aopdata)

# A dataframe, with no spatial geometry
cur <- read_access(city = 'Curitiba', mode = 'public_transport', year = 2019, geometry = FALSE)
```

Let's check the results:

```{r}
dplyr::glimpse(cur )
```

As you can see, a lot of data has been returned from `read_access()`. The dataframe's
columns can be classified into 4 groups, according to the data they contain: _geographic_,
_sociodemographic_, _land use_, and _accessibility_. The following section explains 
the contents of each column.

## Data dictionary

### **Geographic** variables

|**column**|**Description**|**Value**|
|-----|-----|
| `abbrev_muni`|	Abbreviation of city name (3 letters)	|
| `name_muni`  | City name	|
| `code_muni`	 | 7-digit code of each city	|
| `id_hex`	   | Unique id of hexagonal cell	|

### **Sociodemographic** variables

|**column**|**Description**|**Value**|
|-----|-----|
| `P001` | Total number of residents	|
| `P002` | Number of white residents	|
| `P003` | Number of black residents	|
| `P004` | Number of indiginous residents |
| `P005` | Number of asian-descendents residents |
| `R001` | Average household income per capita	| R$ (Brazilian Reais), values in 2010 |
| `R002` | Income quintile group	| 1 (poorest), 2, 3, 4, 5 (richest) |
| `R003` | Income decile group	| 1 (poorest), 2, 3, 4, 5, 6, 7, 8, 9, 10 (richest) |

### **Land use** variables

|**column**|**Description**|**Value**|
|-----|-----|
|	`T001` | Total number of formal jobs |
|	`T002` | Total number of formal jobs with primary education |
|	`T003` | Number of formal jobs with secundary education |
|	`T004` | Number of formal jobs with tertiary education |
|	`E001` | Total number of public schools |
|	`E002` | Number of public schools - early childhood |
|	`E003` | Number of public schools - elementary schools |
|	`E004` | Number of public schools - high schools |
|	`S001` | Total number of healthcare facilities |
|	`S002` | Number of healthcare facilities - low complexity |
|	`S003` | Number of healthcare facilities - medium complexity |
|	`S004` | Number of healthcare facilities - high complexity |


### **Accessibility** variables

The name of the columns with accessibility estimates are the junction of three 
components: 

1) Indicator 

2) Type of opportunity 

3) Time thresold (if applicable)

**1) Indicator**

|**Indicator**|**Description**|**Note**|
|-----|-----|-----|
|`CMA`| Cumulative opportunity measure (active) |  |
|`TMI`| Travel time to closest opportunity | Value = Inf when travel time is longer 
than 2h (public transport) or 1,5h (walking or bicycle) |

**2) Type of opportunity**

|**Indicator**|**Description**|
|-----|-----|
| `TT`	| All jobs |
| `TQ`	| Total jobs with partial match between job education and income quintile |
| `TD`	| Total jobs with partial match between job education and income decile |
| `ST`	| All healthcare facilities |
| `SB`	| Healthcare facilities - Low complexity |
| `SM`	| Healthcare facilities - Medium complexity |
| `SA`	| Healthcare facilities - High complexity |
| `ET`	| All public schools |
| `EI`	| Public schools - early childhood |
| `EF`	| Public schools - elementary schools |
| `EM`	| Public schools - high schools |

**3) Time thresold (only applicable to CMA estimates)**

| **Time thresold**|**Description**|**Note - Only applicable to:**|
|-----|-----|-----|
| `15`| Opportunities accessible within 15 min.	| Active transport modes |
| `30`| Opportunities accessible within 30 min.	| All transport modes |
| `45`| Opportunities accessible within 45 min.	| Active transport modes |
| `60`| Opportunities accessible within 60 min.	| All transport modes |
| `90`| Opportunities accessible within 90 min.	| Public transport |
|`120`| Opportunities accessible within 120 min.| Public transport |

## Next Steps

Now, check the next vignettes for demonstrations on how to use `aopdata` to produce 
[land use](landuse_maps.html) and [accessibility](access_maps.html) maps, as well 
as to analyse [accessibility inequalities](access_inequality.html).





