---
title: "Introduction to `aopdata`"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{Introduction to `aopdata`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette introduces the **`aopdata`** package, lists the cities for which data are available for download, and presents the data dictionary.

**`aopdata`** is an R package that makes it easy to download data from the 
[Access to Opportunities Project (AOP)](https://www.ipea.gov.br/acessooportunidades/en/). 
AOP is a research initiative led by the Institute for Applied Economic Research (Ipea) 
that aims to study transport accessibility and inequalities in access to opportunities 
in Brazilian cities. You can access more information about the project on the [AOP website](https://www.ipea.gov.br/acessooportunidades/en/).


## Installation

You can install `aopdata` from CRAN, or the development version from GitHub.

```{r, message = FALSE, eval = FALSE}
# CRAN
install.packages('aopdata')

# github
devtools::install_github("ipeaGIT/aopdata", subdir = "r-package")
```

## Usage

After installation, you can easily download accessibility estimates, as well as 
population and land use data, from 20 major Brazilian cities. Those cities are:

| **Abbreviation** | **City Name**       |
|--------------|-----------------|
| bel          | Belem           |
| bho          | Belo Horizonte  |
| bsb          | Brasilia        |
| cam          | Campinas        |
| cgr          | Campo Grande    |
| cur          | Curitiba        |
| duq          | Duque de Caxias |
| for          | Fortaleza       |
| goi          | Goiania         |
| gua          | Guarulhos       |
| mac          | Maceio          |
| man          | Manaus          |
| nat          | Natal           |
| poa          | Porto Alegre    |
| rec          | Recife          |
| rio          | Rio de Janeiro  |
| sal          | Salvador        |
| sgo          | Sao Gonçalo     |
| slz          | Sao Luis        |
| spo          | Sao Paulo       |

**Read *accessibility* estimates**

To download accessibility estimates, you can use the `read_access()` function. The
example below downloads data on accessibility by walking for the city of Curitiba, 
in Southern Brazil. The function will download data for the year 2019, which is the 
one currently available.

`read_access()` can also download the spatial geometry of each city by setting
`geometry = TRUE`, which can be used to map accessibility levels across the
city. Refer to the [mapping population and land use](landuse_maps.html) and
[mapping urban accessibility](access_maps.html) vignettes for examples of how
to use the spatial geometry information. For now, we'll set `geometry = FALSE` to 
download a dataframe with no spatial information. 

```{r, message = FALSE, eval = FALSE}
# load the aopdata package
library(aopdata)

# A dataframe, with no spatial geometry
cur <- read_access(city = 'Curitiba', 
                   mode = 'public_transport', 
                   year = 2019, 
                   geometry = FALSE,
                   showProgress = FALSE)
```

Let's check the results:

```{r, message = FALSE, eval = FALSE}
dplyr::glimpse(cur)
```

As you can see, a lot of data has been returned from `read_access()`. The dataframe's
columns can be classified into 4 groups, according to the data they contain: _geographic_,
_sociodemographic_, _land use_, and _accessibility_. The following section explains 
the contents of each column.

## Data dictionary

### **Geographic** variables

|**column**|**Description**|**Value**|
|-----|-----|
| `abbrev_muni`|	Abbreviation of city name (3 letters)	|
| `name_muni`  | City name	|
| `code_muni`	 | 7-digit code of each city	|
| `id_hex`	   | Unique id of hexagonal cell	|

### **Sociodemographic** variables

|**column**|**Description**|**Value**|
|-----|-----|
| `P001` | Total number of residents	|
| `P002` | Number of white residents	|
| `P003` | Number of black residents	|
| `P004` | Number of indigenous residents |
| `P005` | Number of asian-descendent residents |
| `P010` | Number of residents with age between 0 and 5 
| `P011` | Number of residents with age between 6 and 14 
| `P012` | Number of residents with age between 15 and 18
| `P013` | Number of residents with age between 19 and 24
| `P014` | Number of residents with age between 25 and 39
| `P015` | Number of residents with age between 40 and 69
| `P016` | Number of residents with age 70 or more
| `R001` | Average household income per capita	| R$ (Brazilian Reais), values in 2010 |
| `R002` | Income quintile group	| 1 (poorest), 2, 3, 4, 5 (richest) |
| `R003` | Income decile group	| 1 (poorest), 2, 3, 4, 5, 6, 7, 8, 9, 10 (richest) |

### **Land use** variables

|**column**|**Description**|**Value**|
|-----|-----|
|	`T001` | Total number of formal jobs |
|	`T002` | Number of formal jobs with primary education or lower|
|	`T003` | Number of formal jobs with secondary education |
|	`T004` | Number of formal jobs with tertiary education |
|	`E001` | Total number of public schools |
|	`E002` | Number of public schools - early childhood |
|	`E003` | Number of public schools - elementary schools |
|	`E004` | Number of public schools - high schools |
| `M001` | Total number of kids enrolled in public schools |
| `M002` | Number of kids enrolled  in public schools - early childhood |
| `M003` | Number of kids enrolled  in public schools - elementary schools |
| `M004` | Number of kids enrolled  in public schools - high schools |
|	`S001` | Total number of healthcare facilities |
|	`S002` | Number of healthcare facilities - low complexity |
|	`S003` | Number of healthcare facilities - medium complexity |
|	`S004` | Number of healthcare facilities - high complexity |
|	`C001` | Number of centers for social assistance (CRAS) |


### **Accessibility** variables

The name of the columns with accessibility estimates are the junction of three 
components: 

1) Indicator 

2) Type of opportunity 

3) Time thresold (if applicable)

**1) Indicator**

|**Indicator**|**Description**|**Note**|
|-----|-----|-----|
|`CMA`| Active cumulative opportunity measure | The number of opportunities one can access from a given origin  |
|`CMP`| Passive cumulative opportunity measure | The number of people that can access  a given destination |
|`TMI`| Travel time to closest opportunity | Value = Inf when travel time is longer than 2h (public transport) or 1,5h (walking or bicycle) |

**2) Type of opportunity**

|**Indicator**|**Description**|
|-----|-----|
| `TT`	| All jobs |
| `TB`	| Jobs - primary education or lower |
| `TM`	| Jobs - secondary education |
| `TA`	| Jobs - tertiary education |
| `ST`	| All healthcare facilities |
| `SB`	| Healthcare facilities - Low complexity |
| `SM`	| Healthcare facilities - Medium complexity |
| `SA`	| Healthcare facilities - High complexity |
| `ET`	| All public schools |
| `EI`	| Public schools - early childhood |
| `EF`	| Public schools - elementary schools |
| `EM`	| Public schools - high schools |
| `MT`	| Kids enrolled in all public schools |
| `MI`	| Kids enrolled in early childhood |
| `MF`	| Kids enrolled in elementary schools |
| `MM`	| Kids enrolled in high schools |
| `CT`	| Centers for social assistance (CRAS) |


**3) Time thresold (only applicable to CMA estimates)**

| **Time thresold**|**Description**|**Note - Only applicable to:**|
|-----|-----|-----|
| `15`| Opportunities accessible within 15 min.	| All transport modes |
| `30`| Opportunities accessible within 30 min.	| All transport modes |
| `45`| Opportunities accessible within 45 min.	| Active transport modes |
| `60`| Opportunities accessible within 60 min.	| All transport modes |
| `90`| Opportunities accessible within 90 min.	| Public transport and car |
|`120`| Opportunities accessible within 120 min.| Public transport and car |

## Next Steps

Now, check the next vignettes for demonstrations on how to use `aopdata` to produce 
[land use](landuse_maps.html) and [accessibility](access_maps.html) maps, as well 
as to analzse [accessibility inequalities](access_inequality.html).





