---
title: 'Mapping land use data'
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{Mapping land use data} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = identical(tolower(Sys.getenv("NOT_CRAN")), "true"),
  out.width = "100%"
)
```

Here are a few quick examples to illustrate how you can use the {aopdata} package
to map the spatial distribution of activities and urban services in Brazilian cities.


```{r, message = FALSE, warning=FALSE}
# load libraries
library(aopdata)
library(data.table)
library(ggplot2)
library(sf)
library(scales)
```

## Download land use data
```{r, message = FALSE}
df <- aopdata::read_landuse(
  city = 'Fortaleza',
  year = 2019,
  geometry = T,
  showProgress = F
)
```


## Spatial distribution of jobs


```{r, message = FALSE, eval = TRUE}
ggplot() +
  geom_sf(data=df, aes(fill=T001), color=NA, alpha=.9) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  labs(title='Spatial distribution of jobs', fill="N. of jobs") +
  theme_void()

```

```{r, message = FALSE, echo = TRUE}
if (!is.null(df)) {
  ggplot() +
    geom_sf(data=df, aes(fill=T001), color=NA, alpha=.9) +
    scale_fill_distiller(palette = "YlOrRd", direction = 1) +
    labs(title='Spatial distribution of jobs', fill="N. of jobs") +
    theme_void()
}
```


## Spatial distribution of schools

In this case below, elementary schools with the column`E003`.

```{r, message = FALSE, eval = FALSE}
ggplot() +
  geom_sf(data=df, aes(fill=factor(E003)), color=NA, alpha=.9) +
  scale_fill_brewer(palette = "PuBuGn", direction = 1) +
  labs(title='Spatial distribution of elementary schools', fill="N. of schools") +
  theme_void()

```
```{r, message = FALSE, echo = FALSE}
if (!is.null(df)) {  
  ggplot() +
    geom_sf(data=df, aes(fill=factor(E003)), color=NA, alpha=.9) +
    scale_fill_brewer(palette = "PuBuGn", direction = 1) +
    labs(title='Spatial distribution of elementary schools', fill="N. of schools") +
    theme_void()
}
```

## Spatial distribution of healthcare

In this example, we mape high-complexity health care facilities (column `S004`).

```{r, message = FALSE, eval = FALSE}
ggplot() +
  geom_sf(data=df, aes(fill=factor(S004)), color=NA, alpha=.9) +
  scale_fill_brewer(palette = "YlGnBu", direction = 1)+
  labs(title='Spatial distribution of hospitals', fill="N. of hospitals") +
  theme_void()
```

```{r, message = FALSE, echo = FALSE}
if (!is.null(df)) {  
  ggplot() +
    geom_sf(data=df, aes(fill=factor(S004)), color=NA, alpha=.9) +
    scale_fill_brewer(palette = "YlGnBu", direction = 1)+
    labs(title='Spatial distribution of hospitals', fill="N. of hospitals") +
    theme_void()
}
```

## Map Centers for social assistance (CRAS)

```{r, message = FALSE, eval = FALSE}
ggplot() +
  geom_sf(data=df, aes(fill=factor(C001)), color=NA, alpha=.9) +
  scale_fill_brewer(palette = "RdPu", direction = 1)+
  labs(title='Spatial distribution of CRAS facilities', fill="N. of CRAS") +
  theme_void()
```
```{r, message = FALSE, echo = FALSE}
if (!is.null(df)) {  
  ggplot() +
    geom_sf(data=df, aes(fill=factor(C001)), color=NA, alpha=.9) +
    scale_fill_brewer(palette = "RdPu", direction = 1)+
    labs(title='Spatial distribution of CRAS facilities', fill="N. of CRAS") +
    theme_void()
}
```
